### 2017-12-10 readme.rmd
light edits

In a recent paper, [Feuda *et al.* 2017](https://doi.org/10.1016/j.cub.2017.11.008) claim to present strong significant support for Porifera-sister and to have resolved the question at last. Feuda *et al.* present a well-blocked set of phylogenetic analyses that examine support for Ctenophora-sister and Porifera-sister under a variety of conditions. They consider several factors:

@@

Here are the highlights that accompany their paper:

> Highlights
>
> This is the first comparison of model adequacy in non- bilaterian phylogenomics
>
> Animal relationships depend on the ability of the models to describe the data
>
> Sponges are the sister group of the remaining animals
>
> Trees showing comb jellies as the sister of all the other
animals are artifactual

@@

I reorganized the data in this table and color coded the cells to get a better view the general pattern of support for Porifera-sister. This reorganized table is presented below as Figure 2.

@@

*Figure 2 - Posterior probabilities for Porifera-sister. Recodings are: None- original amino acid data, K - KGB-6, D - Dayhoff-6, and S - SR-6.*

@@

The strong conclusions the authors made based largely on Table 3 are therefore rest entirely on recoding the Whelan matrix and applying the CAT-GTR+G model. Even then, less than half (4/9) of these recoded Whelan CAT-GTR+G analyses (Figure 3, right panel, bottom right 9 cells) provide significant support for Porifera-sister.

@@

I am optimistic that there will be significant progress on this problem in coming years. Analysis methods and tools are rapidly improving. New high-quality genome data will soon be available for many more species that are highly relevant to this question. For example, just a few days after Feuda *et al.* 2017 was published, [Belahbib et al.](https://www.biorxiv.org/content/early/2017/12/04/228452) posted a preprint presenting two new sponge genome sequences and analyses directly relevant to interpreting the biological significance of Ctenophora-sister and Porifera-sister. I look forward to to seeing how further advances in analyses and sampling impact tests of Ctenophora-sister and Porifera-sister and their impact on our understanding of animal evolution.

>>>>

In a recent paper, [Feuda *et al.* 2017](https://doi.org/10.1016/j.cub.2017.11.008) claim to present strong, significant support for Porifera-sister and to have resolved the question at last. Feuda *et al.* present a well-blocked set of phylogenetic analyses that examine support for Ctenophora-sister and Porifera-sister under a variety of conditions. They consider several factors:

@@

Here are a couple of the highlights that accompany their paper:

> Sponges are the sister group of the remaining animals
>
> Trees showing comb jellies as the sister of all the other
animals are artifactual

@@

I reorganized the data in this table and color coded the cells to get a better view of the general pattern of support for Porifera-sister. This reorganized table is presented below as Figure 2.

@@

*Figure 2 - Posterior probabilities for Porifera-sister. Recodings are: None- original amino acid data, K - KGB-6, D - Dayhoff-6, and SR-6.*

@@

The strong conclusions the authors made based largely on Table 3 therefore rest entirely on recoding the Whelan matrix and applying the CAT-GTR+G model. Even then, less than half (4/9) of these recoded Whelan CAT-GTR+G analyses (Figure 3, right panel, bottom right 9 cells) provide significant support for Porifera-sister.

@@

I am optimistic that there will be significant progress on this problem in coming years. Analysis methods and tools are rapidly improving. New high-quality genome data will soon be available for many more species that are highly relevant to this question. For example, just a few days after Feuda *et al.* 2017 was published, [Belahbib *et al.*](https://www.biorxiv.org/content/early/2017/12/04/228452) posted a preprint presenting two new sponge genome sequences and analyses directly relevant to interpreting the biological significance of Ctenophora-sister and Porifera-sister. I look forward to to seeing how further advances in analyses and sampling impact tests of Ctenophora-sister and Porifera-sister and their impact on our understanding of animal evolution.
